Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-12-19 Ben_Manuscripts/structure_paper/Draft.tex
Added references, new solvat

  Current commercial RO and NF membranes suffer from inherent limitations of
  their fabrication. Although scalable, each process has a degree of
  stochasticity which makes overcoming the permeability-selectivity tradeoff a
  challenge. Namely, it is difficult to increase the permeability of a desired
  species, while maintaining the same retention of an undesired species.

@@

  Commercial RO membranes are thin film composite membranes with a porous
  support layer and an active layer made of a polymer matrix formed by
  interfacial polymerization. During synthesis, a step-growth polymerization
  reaction occurs at the interface of an aqueous monomer solution, which has been
  soaked into the support layer, and an organic solution of a second monomer. The
  polymer matrix is dense and highly crosslinked which gives high rejections but
  low permeability.

@@

%  solution.  Through the use of a non-solvent, a polymer-rich and polymer-poor
%  phase is formed. A membrane is created as the polymer-poor phase forms pores in
%  the polymer-rich phase. 
%  During phase inversion, a thin polymer film is precipitated out of a polymer
  NF membranes are typically made by a phase inversion process. The most widely
  used phase inversion process is immersion precipitation, during which a polymer
  dissolved in a solvent is submerged in a non-solvent. A solid, porous membrane 
  is all that remains once all solvent has been removed by non-solvent exchange. 
  Resultant pores are polydisperse in size which hurts membrane selectivity.  A
  second technique used to create NF membranes is call track etching in which a
  polymer film is bombarded with charged particles, then chemically etched to
  create pores. The pores are uniform which benefits selectivity, however the
  membranes have a low porosity and subsequent low permeability.

@@

  Development of nanostructured materials has been limited by the ability to
  synthesize and scale various fundamentally sound technologies. Graphene sheets
  are atomically thick which results in excellent water permeability but defects
  during manufacturing severely impact selectivity
  \cite{cohen-tanugi_multilayer_2016}. Molecular dynamics (MD) simulations of carbon
  nanotubes show promise \cite{humplik_nanostructured_2011} but synthetic
  techniques are unable to achieve scalable alignment and pore monodispersity.
  \cite{hata_water-assisted_2004,maruyama_growth_2005} Zeolites have sub-nm pores
  with a narrow pore size distribution and MD simulations exhibit complete
  rejection of solvated ions, \cite{murad_molecular_1998} however, experimental
  rejection was low and attributed to interstitial defects formed during membrane
  synthesis. \cite{li_desalination_2004} There is a need for a scalable
  nanostructured membrane.

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production.  \cite{feng_scalable_2014} LLC systems created by the monomer
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied
  experimentally.
  \cite{smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h} phase
  (Fig.\ref{fig:assembly}). The presence of ca. 10\% wt water results in the
  H\textsubscript{II} phase. In both cases, monomers assemble into mesophases
  made of hexagonally packed, uniform size, cylinders with hydrophilic head
  groups oriented inward towards the cylinder center.  The hydrophilic region can
  act as a pore for aqueous separations \cite{zhou_supported_2005}.  One can
  envision tailoring the pore region for specific separations by changing the
  monomer chemistry \cite{resel_h2-phase_2000}.

  Research into LLC membranes has been revived in recent years. During early
  stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al.~showed that
  the mesophases could be aligned using a magnetic field with subsequent
  crosslinking to lock the structure in place \cite{feng_scalable_2014}. In 2016,
  Feng et al.~showed that the same result could be obtained using a technique
  termed soft confinement \cite{feng_thin_2016}.

@@

  %BJC: TODO: Scalable vector graphic for monomer structure, make bigger. Convert .svg to .eps
  %           Thicken slabs
  %           Make pore region on slab (blue) bigger
  %           Bump up atom sizes in atomistic render
  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.png}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.png}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=\textwidth]{wedge.png}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{layer_blender.png}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{hexagonal_packing.png}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{The liquid crystal monomer Na-GA3C11 (a) rendered atomistically (b)
	exhibits wedge-like character (c). Monomer wedges assemble into disks (d) with
	hydrophilic head groups (blue) facing towards the disk center. The disks
	assemble into hexagonally packed columnar mesophases (e)}~\label{fig:assembly}
  \end{figure}

@@

  A molecular level understanding of LLC membrane structure, enabled by
  molecular dynamics simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  with such an atomistic model, We can directly observe solute transport and
  suggest governing mechanisms. We can observe how the choice of head group may
  influence pore size for size exclusion driven separations. We can interchange
  counterions which may influence both the pore size and the strength of the
  Donnan potential which affects the degree to which the membrane can exclude
  charged species.

@@

  %BJC: I think the first sentence of the following paragraph is
  %necessary to put somewhere, but maybe not in this spot exactly
  In order to appropriately model transport, we must first gain a thorough
  understanding of the nanoscopic structure of LLC membranes. Our approach to
  constructing a general model will follow the development of a model of the
  assembly formed by Na-GA3C11 since it has sufficient experimental
  characterization. We have also narrowed our scope to the development of a model
  of the Col\textsubscript{h} phase membrane. Compared to the H\textsubscript{II}
  phase, the Col\textsubscript{h} phase is a simpler starting point, due to the
  absence of water, and has detailed experimental wide-angle X-ray scattering
  (WAXS) patterns useful for reconstructing structural data. The two
  phases appear to share similar structural characteristics since the pore 
  spacings in each system are in close agreement \cite{resel_structural_2000,feng_thin_2016}.

  Despite having structural data, there is still information which experiment
  cannot definitively answer. Our current level of understanding suggests that
  monomers partition into layers with head groups facing the center, then 
  stack on top of each other to create pores that pack together hexagonally.
  There are several key questions that we will investigate in order to
  expand our understanding.

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self-assembly
  in this system \cite{gazit_possible_2002}. Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations versus a sandwiched
  conformation ~\cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
  \cite{waller_hybrid_2006,ringer_effect_2006}. In this study, we compare
  simulated X-ray diffraction patterns to experiment in order to deduce which
  stacking configurations is most likely.

@@

  	\item A weak intensity line is located at exactly half the $q_z$ value of
	R-$\pi$ ($q_z$ = 0.85 \angstrom$^{-1}$), corresponding to a real space periodic
	spacing of 7.4 \angstrom~. This reflection has been interpreted as
	2\textsubscript{1} helical ordering of aromatic rings along the z axis, meaning
	that if the positions of the aromatic rings can be traced by a helix, then for
	each full turn in the helix, one will encounter two aromatic rings. For this
	reason it will be referred to as R-helix.

	\item A third major reflection is marked by a low intensity ring
	located at r = 1.4 \angstrom$^{-1}$. The real space separation corresponds to
	4.5 \angstrom~ which is characteristic of the average spacing between packed
	alkane chains.  This reflection will be called R-alkanes.

	\item Within R-alkanes, are four spots of higher relative intensity
	which will be called R-spots. All are located $\sim 37$ degrees from the $q_z$
	axis in their respective quadrants. In many liquid crystal systems this can be
	explained by the tilt angle of the alkane chains with respect to the membrane
	plane.

	\item The final feature corresponds to the spacing and symmetry of the
	d\textsubscript{100} plane which can be related to the distance between pores.
	The feature, which will be called R-pores, is characterized by dots along $q_z$
	= 0. The spacing between dots is indicative of the hexagonal symmetry of the
	packed pores. The same information at higher resolution is obtained using SAXS.

@@

  %BJC: TODO: update colorbar
  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
%MRS3: would be good to have all feature labels agree: either can edit to put the names on, or have a key relating numbers to the names you use.
%BJC2: I will annotate it with the names R-pi, R-alkanes etc. Is that what you mean?
%MRS4: that works fine.  We just need a way to connect what is in the figure with the names used in the text.  Ideally there would not be two sets of labels, just one.
%BJC3: Got it, will do once I have more time to focus on figures 
	\caption{(a) The repeat spacing in the 1D small angle X-ray scattering pattern
		is characteristic of hexagonal packing. The leading peak represents the
		distance between the d\textsubscript{100} planes. Using this distance, we know
		that the distance between pore centers is 4.12 nm. (b) 2D wide angle X-ray
		scattering gives details about repeating features less than 1 nanometer
		apart. Experimentalists have justified each of the 5 major reflections
		present as follows: (I) Aromatic head group $\pi-\pi$ stack 3.7 \AA~apart.
		(II) Monomers arrange vertically in a 2\textsubscript{1} helix. (III) Alkane
 		chain tails pack 4.5 \AA~apart. (IV) Monomer tails tilt with respect to the 
 		membrane plane. (V) As derived from SAXS, pores pack hexagonally and are 
		spaced 4.12 nm apart}
	\label{fig:SWAXS}
 \end{figure}

@@

  Liquid crystal monomers were parameterized using the Generalized AMBER
  Forcefield \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. Atomic charges were assigned using the am1bccsym
  method of \texttt{molcharge} shipped with QUACPAC from Openeye Scientific
  Software. All molecular dynamics simulations were run using Gromacs 2016.
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}
  the experiments.  An ensemble of characteristic, low-energy vacuum monomer
  configurations were constructed by applying a simulated annealing process to a
  parameterized monomer. Monomers were cooled from 1000K to 50K over 10
  nanoseconds.  A low energy configuration was randomly pulled from the
  trajectory and charges were reassigned using \texttt{molcharge}.  Using the new
  charges, the monomer system was annealed again and a random monomer
  configuration was pulled from the trajectory to be used for full system
  construction (Figure~\ref{fig:python}a).  See supplemental information for
  further detail.

@@

  %BJC: TODO: reproduce figure in latex (or gimp?) with higher resolution arrows
  \begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{build.PNG} %BJC: put an xyz axis with the unit cell
	\caption{(a) A single monomer was parameterized and annealed to produce a low energy
		configuration. (b) Monomers are rotated and assembled into layers with 
		hydrophlic centers. (c) Twenty layers are stacked on top of each other to create
		a pore. (d) Pores are duplicated and placed into a monoclinic unit cell}\label{fig:python}
  \end{figure}

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of twenty stacked monomer layers with periodic
  continuity in the z direction, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. A small number of layers is
  preferred in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomer
  layers in each pore in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. %This point will be explained in more detail later.
%MRS3: isn't there some observations you need at least 10?  
%BJC2: When I first set up the system, there was a gap and a small number of layers 
% appeared to cause the phase to disassociate. However, I can get a pretty small
% number of layers in the infinite pore system, but the diffraction is low resolution.

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure when a
  reasonable value is chosen (See Supplemental Information). The pore radius is
  chosen to be 0.6 nm in our initial configurations because the pore size is
  estimated to be $\approx$ 1.2 nm \cite{zhou_supported_2005}.  The initial
  distance between pores also has little effect on the the equilibrated
  structure. However, one should not start them too close or there will be high
  energy repulsions during early equilibration. We chose an initial pore spacing
  of 4.5 nm, $\sim$10\% larger than the experimental value of 4.12 nm.  The
  distance between layers, the rotation of the layers with respect to adjacent
  layers, and the number of monomers per layer do influence the equilibrium
  structure and require further justification for their choices.  We rely on
  experimental data to inform them.

@@

  The relative interlayer orientation was chosen based on clues from
  diffraction data as well as the various known stacking modes of benzene and
  substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  ~\cite{sinnokrot_estimates_2002} (\Cref{fig:sandwiched,fig:pd,fig:tshaped}).
  The T-shaped configuration was ruled out because its $\sim$5 \angstrom~
  equilibrium stacking distance ~\cite{sinnokrot_estimates_2002} is inconsistent
  with R-$\pi$. It is also infeasible for the monomers to orient in the T-shaped
  conformation because of the bulky tail groups. The system's preference towards
  the sandwiched vs. parallel displaced stacking modes will be explored in some
  detail.  Both have reported stacking distances near the R-$\pi$ value of 3.7
  \angstrom. Headgroups in our sandwiched initial configuration are stacked
  directly on top of each other while stacked headgroups in the parallel
  displaced initial configuration are offset by $180/nmon$ degrees where $nmon$
  equals the number of monomers per layer.

  The number of monomers in each layer is unknown, as state in
  (\ref{point:monomernum}). We tested configurations constructed with a varied
  number of monomers per layer. Systems were built in the offset and parallel
  displaced configurations with 4, 5, 6, 7 and 8 monomers per layer.

@@

  In order to fully match synthetic procedures, we created a crosslinking
  algorithm that can be applied to equilibrated structures. The purpose of
  crosslinking is to maintain macroscopic alignment of the crystalline domains,
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data. The
  algorithm was developed based on the known reaction mechanism. Crosslinking of
  this system is a free radical polymerization (FRP) taking place between
  terminal vinyl groups present on each of the three monomer tails. FRPs require
  an initiator which bonds to the system, meaning new atoms are introduced into
  the system. For simplicity, the initiator was simulated as hydrogen and made
  present in the simulation by including them in all possible locations where an
  addition could occur as dummy atoms. The crosslinking procedure is carried out
  iteratively. During each iteration, bonding carbon atoms are chosen based on a
  distance cut-off. The topology is updated with new bonds and dummy hydrogen
  atoms are changed to appropriate hydrogen types. Head-to-tail addition was the
  only propagation mode considered due to its dominance in the real system.
  Direction of attack was not considered because the resultant mixture is
  racemic.

  Our implementation requires long simulation times to achieve high degrees of
  crosslinking. A typical crosslinking procedure can take up to 24 hours. In
  order to collect equilibrated data, further NPT simulation is necessary. We
  typically run a crosslinked system for an additional 100 ns to allow the system
  to readjust. For those reasons we did not crosslink all systems tested, but only
  the most promising structure. We show that crosslinking does not significantly
  change any of our drawn conclusions in Section 3.6.

@@

  In systems built with 5\% wt water, the pore region becomes filled with
  water. We plotted the number density of components in this system. As with the 
  dry systems, we see a gradual compositional transition from hydrophilic to hydrophobic.  
  We see that the pores become a mixture of water molecules and sodium ions (Fig.
  \ref{fig:water_density}).

>>>>

  Current commercial RO and NF membranes suffer limitations inherent to their
  fabrication. Although scalable, each process has a degree of stochasticity
  which makes overcoming the permeability-selectivity tradeoff a challenge.
  Namely, it is difficult to increase the permeability of a desired species,
  while maintaining the same retention of an undesired species
  \cite{werber_materials_2016}.

@@

  Commercial RO membranes are typically thin film composite membranes with a
  porous support layer and an active layer made of a polymer matrix formed
  through interfacial polymerization \cite{jeong_interfacial_2007}. During
  interfacial polymerization, a step-growth polymerization reaction occurs at the
  interface of an aqueous monomer solution, which has been soaked into the
  support layer, and an organic solution of a second monomer. The polymer matrix
  is dense and highly crosslinked which gives high rejections but low
  permeability.

@@

%  solution.  Through the use of a non-solvent, a polymer-rich and polymer-poor
%  phase is formed. A membrane is created as the polymer-poor phase forms pores in
%  the polymer-rich phase. 
%  During phase inversion, a thin polymer film is precipitated out of a polymer
  NF membranes are typically made by a phase inversion process  The most widely
  used phase inversion process is immersion precipitation, during which a polymer
  dissolved in a solvent is submerged in a non-solvent. A solid, porous membrane
  is all that remains once all solvent has been removed by non-solvent exchange
  \cite{smolders_microstructures_1992}.  Resultant pores are polydisperse in size
  which hurts membrane selectivity. A second technique used to create NF
  membranes is call track etching in which a polymer film is bombarded with
  charged particles, then chemically etched to create pores
  \cite{apel_track_2001}. The pores are uniform which benefits selectivity,
  however the membranes have a low porosity and subsequently low permeability.

@@

  Development of nanostructured materials has been limited by the ability to
  synthesize and scale various fundamentally sound technologies. Graphene sheets
  are atomically thick which results in excellent water permeability but defects
  during manufacturing severely impact selectivity
  \cite{cohen-tanugi_multilayer_2016}. Molecular dynamics (MD) simulations of carbon
  nanotubes show promise \cite{humplik_nanostructured_2011} but synthetic
  techniques are unable to achieve scalable alignment and pore monodispersity
  \cite{hata_water-assisted_2004,maruyama_growth_2005}. Zeolites have sub-nm pores
  with a narrow pore size distribution and MD simulations exhibit complete
  rejection of solvated ions \cite{murad_molecular_1998}, however, experimental
  rejection was low and attributed to interstitial defects formed during membrane
  synthesis \cite{li_desalination_2004}. There is a need for a scalable
  nanostructured membrane.

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production \cite{feng_scalable_2014}. LLC systems created by the monomer
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied
  experimentally
  \cite{feng_scalable_2014,smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_thin_2016}.
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h} phase
  (Fig.\ref{fig:assembly}). The presence of ca. 10\% wt water results in the
  H\textsubscript{II} phase. In both cases, monomers assemble into mesophases
  made of hexagonally packed, uniform size, cylinders with hydrophilic head
  groups oriented inward towards the cylinder center.  The hydrophilic region can
  act as a pore for aqueous separations \cite{zhou_supported_2005}.  One can
  envision tailoring the pore region for specific separations by changing the
  monomer chemistry \cite{resel_h2-phase_2000}.

  Research into LLC membranes has been revived in recent years. During early
  stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al.~showed that
  the mesophases could be aligned using a magnetic field with subsequent
  crosslinking to lock the structure in place \cite{feng_scalable_2014}. In 2016,
  Feng et al. showed that the same result could be obtained using a second technique
  termed soft confinement \cite{feng_thin_2016}.

@@

  %BJC: TODO: Scalable vector graphic for monomer structure, make bigger. Convert .svg to .eps
  %           Thicken slabs
  %           Make pore region on slab (blue) bigger
  %           Bump up atom sizes in atomistic render
  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.png}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.png}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=\textwidth]{wedge.png}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{layer_blender.png}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{hexagonal_packing.png}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{The liquid crystal monomer Na-GA3C11 (a) rendered atomistically (b)
	exhibits wedge-like character (c). Monomer wedges assemble into disks (d) with
	hydrophilic head groups (blue) facing towards the disk center. The disks
	assemble into hexagonally packed columnar mesophases (e).}~\label{fig:assembly}
  \end{figure}

@@

  A molecular level understanding of LLC membrane structure, enabled by
  molecular dynamics simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  With such an atomistic model, we can directly observe solute transport and
  suggest governing mechanisms. We can observe how the choice of head group may
  influence pore size for size exclusion driven separations. We can interchange
  counterions which may influence both the pore size and the strength of the
  Donnan potential which affects the degree to which the membrane can exclude
  charged species.

@@

  %BJC: I think the first sentence of the following paragraph is
  %necessary to put somewhere, but maybe not in this spot exactly
  In order to appropriately model transport, we must first gain a thorough
  understanding of the nanoscopic structure of LLC membranes. Our approach to
  constructing a general model will follow the development of a model of the
  assembly formed by Na-GA3C11 since it has sufficient experimental
  characterization. We have also narrowed our scope to the development of a model
  of the Col\textsubscript{h} phase membrane. Compared to the H\textsubscript{II}
  phase, the Col\textsubscript{h} phase is a simpler starting point, due to the
  absence of water, and has detailed experimental wide-angle X-ray scattering
  (WAXS) patterns useful for reconstructing structural data. The two
  phases appear to share similar structural characteristics since the pore 
  spacings in each system are in close agreement \cite{feng_thin_2016,resel_structural_2000}.

  Despite having structural data, there is still information which experiment
  cannot definitively answer. Our current level of understanding suggests that
  monomers partition into layers with head groups facing towards the center, then 
  stack on top of each other to create pores that pack together hexagonally.
  There are several key questions that we will investigate in order to
  expand this understanding:

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self-assembly
  in this system \cite{gazit_possible_2002}. Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations versus a sandwiched
  conformation \cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
  \cite{waller_hybrid_2006,ringer_effect_2006} In this study, we compare
  simulated X-ray diffraction patterns to experiment in order to deduce which
  stacking configuration is most likely.

@@

	\item A weak intensity line is located at exactly half the $q_z$ value
	of R-$\pi$ ($q_z$ = 0.85 \angstrom$^{-1}$), corresponding to a real space
	periodic spacing of 7.4 \angstrom~. This reflection has been interpreted as
	2\textsubscript{1} helical ordering of aromatic rings along the z axis
	\cite{feng_scalable_2014}, meaning that if the positions of the aromatic rings
	can be traced by a helix, then for each full turn in the helix, one will
	encounter two aromatic rings. For this reason it will be referred to as
	R-helix.

	\item A third major reflection is marked by a low intensity ring
	located at r = 1.4 \angstrom$^{-1}$. The real space separation corresponds to
	4.5 \angstrom~ which is characteristic of the average spacing between packed
	alkane chains \cite{mcintosh_organization_1980}. This reflection will be called
	R-alkanes.

	\item Within R-alkanes, are four spots of higher relative intensity
	which will be called R-spots. All are located $\sim 37$ degrees from the $q_z$
	axis in their respective quadrants. In many liquid crystal systems this can be
	explained by the tilt angle of the alkane chains with respect to the membrane
	plane \cite{govind_simple_2001}.

	\item The final feature corresponds to the spacing and symmetry of the
	d\textsubscript{100} plane which can be related to the distance between pores.
	The feature, which will be called R-pores, is characterized by dots along $q_z$
	= 0. The spacing between dots is indicative of the hexagonal symmetry of the
	packed pores. The same information at higher resolution is obtained using SAXS (Fig.~\ref{fig:SAXS}).

@@

  %BJC: TODO: update colorbar
  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
%MRS3: would be good to have all feature labels agree: either can edit to put the names on, or have a key relating numbers to the names you use.
%BJC2: I will annotate it with the names R-pi, R-alkanes etc. Is that what you mean?
%MRS4: that works fine.  We just need a way to connect what is in the figure with the names used in the text.  Ideally there would not be two sets of labels, just one.
%BJC3: Got it, will do once I have more time to focus on figures 
	\caption{(a) (Reproduced from \cite{feng_thin_2016}) The repeat spacing
		in the 1D small angle X-ray scattering pattern is characteristic of hexagonal
		packing. The leading peak represents the distance between the
		d\textsubscript{100} planes. Using this distance, we know that the distance
		between pore centers is 4.12 nm. (b) 2D wide angle X-ray scattering gives
		details about repeating features less than 1 nanometer apart. Experimentalists
		have justified each of the 5 major reflections present as follows: (I) Aromatic
		head group $\pi-\pi$ stack 3.7 \AA~apart.  (II) Monomers arrange vertically in
		a 2\textsubscript{1} helix. (III) Alkane chain tails pack 4.5 \AA~apart. (IV)
		Monomer tails tilt with respect to the membrane plane. (V) As derived from
		SAXS, pores pack hexagonally and are spaced 4.12 nm apart}
	\label{fig:SWAXS}
 \end{figure}

@@

  Liquid crystal monomers were parameterized using the Generalized AMBER
  Forcefield \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. Atomic charges were assigned using the am1bccsym
  method of \texttt{molcharge} shipped with QUACPAC from Openeye Scientific
  Software. All molecular dynamics simulations were run using Gromacs 2016
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}.
  An ensemble of characteristic, low-energy vacuum monomer
  configurations were constructed by applying a simulated annealing process to a
  parameterized monomer. Monomers were cooled from 1000K to 50K over 10
  nanoseconds.  A low energy configuration was randomly pulled from the
  trajectory and charges were reassigned using \texttt{molcharge}.  Using the new
  charges, the monomer system was annealed again and a random monomer
  configuration was pulled from the trajectory to be used for full system
  construction (Figure~\ref{fig:python}a).  See supplemental information for
  further detail.

@@

  %BJC: TODO: reproduce figure in latex (or gimp?) with higher resolution arrows
  \begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{build.PNG} %BJC: put an xyz axis with the unit cell
	\caption{(a) A single monomer was parameterized and annealed to produce a low energy
		configuration. (b) Monomers are rotated and assembled into layers with 
		hydrophilic centers. (c) Twenty layers are stacked on top of each other to create
		a pore. (d) Pores are duplicated and placed into a monoclinic unit cell}\label{fig:python}
  \end{figure}

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of twenty stacked monomer layers with periodic
  continuity along the pore axis, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. A small number of layers is
  preferred in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomer
  layers in each pore in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. %This point will be explained in more detail later.
%MRS3: isn't there some observations you need at least 10?  
%BJC2: When I first set up the system, there was a gap and a small number of layers 
% appeared to cause the phase to disassociate. However, I can get a pretty small
% number of layers in the infinite pore system, but the diffraction is low resolution.

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure when a
  reasonable value is chosen (See Supplemental Information). The pore radius is
  chosen to be 0.6 nm in our initial configurations because the pore size is
  estimated to be $\sim$1.2 nm \cite{zhou_supported_2005}. The initial distance
  between pores also has little effect on the the equilibrated structure (See
  supplemental information). However, one should not start them too close or
  there will be high energy repulsions during early equilibration. We chose an
  initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental value
  of 4.12 nm.  The distance between layers, the rotation of the layers with
  respect to adjacent layers, and the number of monomers per layer do influence
  the equilibrium structure and require further justification for their choices.
  We rely on experimental data to inform them.

@@

  The relative interlayer orientation was chosen based on clues from
  diffraction data as well as the various known stacking modes of benzene and
  substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  \cite{sinnokrot_estimates_2002} (Fig. \ref{fig:sandwiched} and \ref{fig:pd,fig:tshaped}).
  The T-shaped configuration was ruled out because its $\sim$5 \angstrom~
  equilibrium stacking distance \cite{sinnokrot_estimates_2002} is inconsistent
  with R-$\pi$. It is also infeasible for the monomers to orient in the T-shaped
  conformation because of the bulky tail groups. The system's preference towards
  the sandwiched vs. parallel displaced stacking modes will be explored in some
  detail. Both have reported stacking distances near the R-$\pi$ value of 3.7
  \angstrom. Headgroups in our sandwiched initial configuration are stacked
  directly on top of each other while stacked headgroups in the parallel
  displaced initial configuration are offset by $180/nmon$ degrees where $nmon$
  equals the number of monomers per layer.

  The number of monomers in each layer is unknown, as stated in
  (\ref{point:monomernum}). We tested configurations constructed with a varied
  number of monomers per layer. Systems were built in the offset and parallel
  displaced configurations with 4, 5, 6, 7 and 8 monomers per layer.

@@

  In order to fully match synthetic procedures, we created a crosslinking
  algorithm that can be applied to equilibrated structures. The purpose of
  crosslinking is to maintain macroscopic alignment of the crystalline domains,
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data. The
  algorithm was developed based on the known reaction mechanism. Crosslinking of
  this system is a free radical polymerization (FRP) taking place between
  terminal vinyl groups present on each of the three monomer tails. FRPs require
  an initiator which bonds to the system, meaning new atoms are introduced into
  the system. For simplicity, the initiator was simulated as hydrogen and made
  present in the simulation by including them in all possible locations where an
  addition could occur as dummy atoms. The crosslinking procedure is carried out
  iteratively. During each iteration, bonding carbon atoms are chosen based on a
  distance cut-off. The topology is updated with new bonds and dummy hydrogen
  atoms are changed to appropriate hydrogen types. Head-to-tail addition was the
  only propagation mode considered due to its dominance in the real system \cite{young_introduction_2011}.
  Direction of attack was not considered because the resultant mixture is
  racemic.

  Our implementation requires long simulation times to achieve high crosslink 
  densities. A typical crosslinking procedure can take up to 24 hours. In
  order to collect equilibrated data, further NPT simulation is necessary. We
  typically run a crosslinked system for an additional 100 ns to allow the system
  to readjust. For those reasons we did not crosslink all systems tested, but only
  the most promising structure. We show that crosslinking does not significantly
  change any of our drawn conclusions in Section 3.6.

@@

  In systems built with 5 wt\% water, the pore region becomes filled with
  water. We plotted the number density of components in this system. As with the 
  dry systems, we see a gradual compositional transition from hydrophilic to hydrophobic.  
  We see that the pores become a mixture of water molecules and sodium ions (Fig.
  \ref{fig:water_density}).
